---
title: "PROJECT_RT_A1_KELOMPOK 4_164231059"
author: "Faiz Iqbal I'tishom"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Library
```{r}
library(readxl)
library(skedastic)
library(lmtest)
library(regclass)
library(ggplot2)
library(reshape2)
library(dplyr)
library(nlme)
library(ggpubr)
library(boot)
library(MASS)        
library(robustbase)  
library(caret)  
library(leaps)
```

```{r}
library(readr)
data4 <- read_csv("data 4.csv")
View(data4)
```
Subset Data
```{r}
y = data4$ISI
x1 = data4$X
x2 = data4$Y
x3 = data4$month
x4 = data4$day
x5 = data4$FFMC
x6 = data4$DMC
x7 = data4$DC
x8 = data4$temp
x9 = data4$RH
x10 = data4$wind
x11 = data4$rain
x12 = data4$area
```

Check Missing Value
```{r}
missing_summary <- sapply(data4, function(x) sum(is.na(x)))
missing_percentage <- sapply(data4, function(x) mean(is.na(x)) * 100)
missing_values <- data.frame(
  Variable = names(missing_summary),
  MissingCount = missing_summary,
  MissingPercentage = missing_percentage
)
print("Summary of Missing Values:")
print(missing_values)
```

Outlier dengan IQR
```{r}
detect_outliers <- function(column) {
  Q1 <- quantile(column, 0.25, na.rm = TRUE)
  Q3 <- quantile(column, 0.75, na.rm = TRUE)
  IQR <- Q3 - Q1
  lower_bound <- Q1 - 1.5 * IQR
  upper_bound <- Q3 + 1.5 * IQR
  sum(column < lower_bound | column > upper_bound, na.rm = TRUE)
}

outlier_summary <- sapply(data4, function(x) {
  if (is.numeric(x)) detect_outliers(x) else NA
})
outliers <- data.frame(
  Variable = names(outlier_summary),
  OutlierCount = outlier_summary
)
print("Summary of Outliers:")
print(outliers)
```

Outlier dengan Boxplot
```{r}
numeric_columns <- sapply(data4, is.numeric)
numeric_data <- data4[, numeric_columns]

boxplot_data <- data.frame(stack(numeric_data))
names(boxplot_data) <- c("Value", "Variable")

ggplot(boxplot_data, aes(x = Variable, y = Value)) +
  geom_boxplot(outlier.color = "red", outlier.shape = 8) +
  theme_minimal() +
  labs(
    title = "Boxplot of Variables with Outliers Highlighted",
    x = "Variables",
    y = "Values"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
Pemilihan Variabel dengan Best Subset
```{r}
data_subset <- data.frame(ISI = y, X = x1, Y = x2, month = x3, day = x4,
                          FFMC = x5, DMC = x6, DC = x7, temp = x8, RH = x9,
                          wind = x10, rain = x11, area = x12)

# Remove rows with missing values (if any)
data_subset <- na.omit(data_subset)

# Perform Best Subset Selection
best_subset <- regsubsets(ISI ~ ., data = data_subset, nvmax = 12)

# Summary of Best Subset Selection
subset_summary <- summary(best_subset)

# Display metrics for each model size
print("Metrics for Best Subset Selection:")
metrics <- data.frame(
  NumVariables = 1:length(subset_summary$cp),
  Adjusted_R2 = subset_summary$adjr2,
  Cp = subset_summary$cp,
  BIC = subset_summary$bic
)
print(metrics)

# Plot Adjusted R2, Cp, and BIC for different model sizes
par(mfrow = c(1, 3))  # Arrange plots in 1 row, 3 columns

# Adjusted R2 plot
plot(metrics$NumVariables, metrics$Adjusted_R2, type = "b", pch = 19,
     xlab = "Number of Variables", ylab = "Adjusted R2", main = "Adjusted R2")

# Cp plot
plot(metrics$NumVariables, metrics$Cp, type = "b", pch = 19,
     xlab = "Number of Variables", ylab = "Cp", main = "Mallows' Cp")

# BIC plot
plot(metrics$NumVariables, metrics$BIC, type = "b", pch = 19,
     xlab = "Number of Variables", ylab = "BIC", main = "Bayesian Information Criterion")

# Highlight the best model
```
```{r}
best_model_index <- which.max(subset_summary$adjr2)
print(paste("Best model with", best_model_index, "variable(s):"))
print(subset_summary$outmat[best_model_index, ])
```

Model OLS
```{r}
ols <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12,  data = data4)
ols
summary(ols)
```
```{r}
ols <- lm(y ~ x3 + x4 + x5 + x7 + x10,  data = data4)
ols
summary(ols)
```

Uji Homoskedastisitas
```{r}
glejser(ols)
```

Uji Autokorelasi
```{r}
dwtest(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, data = data4)
```

Uji Multikolinearitas
```{r}
VIF(ols)
```
Visualisasi Multikolinearitas dengan Heatmap
```{r}
numerik_data <- data4[sapply(data4, is.numeric)]
cor_matrix <- cor(numerik_data, use = "complete.obs")
```

```{r}
cor_data <- melt(cor_matrix)
colnames(cor_data) <- c("Variable1", "Variable2", "Correlation")
```

```{r}
heatmap_plot <- ggplot(cor_data, aes(x = Variable1, y = Variable2, fill = Correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#ffc54a", high = "#821716", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Korelasi") +
  geom_text(aes(label = round(Correlation, 2)), color = "black", size = 3) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Heatmap Korelasi", x = "Variabel", y = "Variabel")

print(heatmap_plot)
```

Uji Normalitas
```{r}
residuals <- resid(ols)

shapiro_test <- shapiro.test(residuals)
print(shapiro_test)

ggqqplot(residuals, title = "Q-Q Plot of Residuals")
```
**Transformasi Variabel**

```{r}
subset_data <- data.frame(
  y, x1, x2, x3, x4, x5, x6, x7, x8, x9, x10, x11, x12
)

# Convert categorical variables ('month' and 'day') to factors
subset_data$x3 <- as.factor(subset_data$x3)
subset_data$x4 <- as.factor(subset_data$x4)

# Apply log transformation to continuous variables
# Avoid log(0) by adding a small constant (e.g., 1e-6) if necessary
subset_data$log_y <- log(subset_data$y + 1e-6)
subset_data$log_x5 <- log(subset_data$x5 + 1e-6)
subset_data$log_x6 <- log(subset_data$x6 + 1e-6)
subset_data$log_x7 <- log(subset_data$x7 + 1e-6)
subset_data$log_x8 <- log(subset_data$x8 + 1e-6)
subset_data$log_x9 <- log(subset_data$x9 + 1e-6)
subset_data$log_x10 <- log(subset_data$x10 + 1e-6)
subset_data$log_x11 <- log(subset_data$x11 + 1e-6)
subset_data$log_x12 <- log(subset_data$x12 + 1e-6)

# Build the OLS regression model with transformed variables
ols_model <- lm(
  log_y ~ x1 + x2 + x3 + x4 + log_x5 + log_x6 + log_x7 + log_x8 + log_x9 + 
          log_x10 + log_x11 + log_x12, 
  data = subset_data
)

# Summarize the model
summary(ols_model)

# Plot diagnostics
par(mfrow = c(2, 2)) # Arrange plots in a 2x2 grid
plot(ols_model)
```

Subset data kembali untuk mempermudah pengujian GLS dan WLS
```{r}
y <- data4$ISI
x1 <- data4$X
x2 <- data4$Y
x3 <- as.factor(data4$month)
x4 <- as.factor(data4$day)
x5 <- data4$FFMC
x6 <- data4$DMC
x7 <- data4$DC
x8 <- data4$temp
x9 <- data4$RH
x10 <- data4$wind
x11 <- data4$rain
x12 <- data4$area

subset_data <- data.frame(y, x1, x2, x3, x4, x5, x6, x7, x8, x9, x10, x11, x12)
```

**Model GLS**
```{r}
model_gls <- gls(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, data = subset_data,
                 method = "REML")  # Method 'REML' adalah default untuk estimasi parameter

# Menampilkan ringkasan model GLS
cat("Ringkasan model GLS:\n")
summary(model_gls)

# Memeriksa struktur residuals untuk mengevaluasi heteroskedastisitas
cat("\nAnalisis residual:\n")
plot(model_gls, which = 1:2)  # Grafik residual dan QQ plot

```

**Model WLS**
```{r}
model_ols <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, data = subset_data)
residuals_ols <- abs(resid(model_ols))
weights <- 1 / (residuals_ols^2) 
model_wls <- lm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, data = subset_data, weights = weights)
summary(model_wls)
```

**Membandingkan AIC dan BIC antara model OLS, GLS, dan WLS**
```{r}
AIC(model_ols, model_gls, model_wls)
BIC(model_ols, model_gls, model_wls)
```

**Regresi Robust**

Estimasi-M
```{r}
EM <- rlm(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, method = "M")
summary(EM)
print(EM)
```

Estimasi-LTS
```{r}
subset_data <- data.frame(y, x1, x2, x3, x4, x5, x6, x7, x8, x9, x10, x11, x12)

subset_data$x3 <- as.factor(subset_data$x3)
subset_data$x4 <- as.factor(subset_data$x4)

lts_model <- ltsReg(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, data = subset_data)

summary(lts_model)
```

Estimasi-S
```{r}
S <- lmrob(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, method = "S")
summary(S)
```

```{r}
VIF(S)
```


Uji Normalitas
```{r}
residuals <- resid(S)

shapiro_test <- shapiro.test(residuals)
print(shapiro_test)

ggqqplot(residuals, title = "Q-Q Plot of Residuals")
```


Estimasi-MM
```{r}
MM <- lmrob(y ~ x1 + x2 + x3 + x4 + x5 + x6 + x7 + x8 + x9 + x10 + x11 + x12, method ="MM")
summary(MM)
```


**Kesimpulan**
